---
title: "Spatial GLMs"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(ggmap)
library(knitr)
library(gtools)
library(spBayes)
library(mgcv)
library(mnormt)
library(arm)
library(rstanarm)
library(rstan)
```

## Generalized Linear Model Notation
There are three components to a generalized linear model:

1. Sampling Distribution: 
\vfill

2. Linear combination of predictors: 
\vfill

3. A link function to 

\vfill

## Binary Regression Overview
Write out the complete model specification for binary regression.
\vfill


\vfill



\newpage
Latent interpretation of probit model:


\vfill


\vfill


Consider air quality data from Colorado as a motivating example.

```{r, echo = T, echo = F}
load('CO_Air.Rdata')

co_box <- c(left = min(CO$Longitude)-.25, bottom = min(CO$Latitude)-.25, right = max(CO$Longitude)+.25, top =max(CO$Latitude)+.25)
co_map <- get_map(co_box) %>% ggmap()
  
#co_map + geom_point(aes(x = Longitude, y = Latitude, color= Exceedance_Count), data = CO, inherit.aes = F) + ggtitle("Ozone Measurements")

co_map + geom_point(aes(x = Longitude, y = Latitude, color= Exceedance), data = CO, inherit.aes = F) + ggtitle("Ozone Measurements")
```

\newpage


Interpret the output.

```{r, echo = T}
CO <- CO %>% mutate(north = as.numeric(Latitude > 38 ))
glm(Exceedance~north, family=binomial(link = 'probit'),data=CO) %>% display()
```

\vfill

```{r, echo = T}
glm(Exceedance~north, family=binomial(link = 'logit'),data=CO) %>% display()
```


\vfill

## Spatial Binary Regression
Assume $Y(\boldsymbol{s_i})$ is the binary response for $\boldsymbol{s_i}$,

\vfill


\newpage

## Simulating spatial random effects for binary data


```{r, eval = T, echo = T}
N.sim <- 100
Lat.sim <- runif(N.sim,37,40)
Long.sim <- runif(N.sim,-109,-104)
phi.sim <- 1
sigmasq.sim <- 1
beta.sim <- c(-1,1)
north.sim <-  as.numeric(Lat.sim > 38)


d <- dist(cbind(Lat.sim,Long.sim), upper = T, diag = T) %>% as.matrix
H.sim <- sigmasq.sim * exp(- d / phi.sim)
w.sim <- rmnorm(1,0,H.sim)
xb.sim <- beta.sim[1] + beta.sim[2] * north.sim
y.sim <- rbinom(N.sim,1,pnorm(xb.sim + w.sim))
```

\newpage

```{r}
tibble(y = Lat.sim, x = Long.sim, response = factor(y.sim)) %>%
  ggplot(aes(y=y, x=x, color = response)) + geom_point() +
  theme_bw() + ggtitle('Binary Response') + scale_color_manual(values=c("#023FA5", "#8E063B"))

tibble(y = Lat.sim, x = Long.sim, `random \neffect` = w.sim) %>%
  ggplot(aes(y=y, x=x, color = `random \neffect`)) + geom_point() +
  theme_bw() + ggtitle('Spatial Random Effect') + 
  scale_color_gradientn(colours = colorspace::diverge_hcl(7))
```

\newpage

#### STAN: probit regression

```{r}
writeLines(readLines('probit_regression.stan'))
```


# Binary Regression
```{r warning = F, message = F, results = 'hide', echo = T}
probit_stan <- stan(file = 'probit_regression.stan', data = list(N = N.sim, y = y.sim, x = north.sim))
```

```{r warning = F, message = F, echo = T}
print(probit_stan, pars = c('beta0', 'beta1'))
glm(y.sim ~ north.sim, family = binomial(link = 'probit'))
tibble(y.sim = y.sim, north.sim = north.sim) %>% stan_glm(y.sim ~ north.sim, family = binomial(link = 'probit'), refresh = 0, data = .)
```



\newpage

# Spatial Poisson Regression

## Motivation

```{r}

co_map + geom_point(aes(x = Longitude, y = Latitude, color= Exceedance_Count), data = CO, inherit.aes = F) + ggtitle("Ozone Measurements")

```


## Poisson Regression Overview
Write out the complete model specification for Poisson regression.

\vfill


\vfill

\vfill


